Supplementary material

### 2022-02-03 submission/manuscript.rmd
new updates to manuscript

```{r rmsebase, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated from a simulated super-population of size 500,000. The scenario with true constant relative treatment effect (panel A) had a true prediction AUC of 0.75 and sample size of 4250. The RMSE is also presented for strong linear (panel B), strong quadratic (panel C), and non-monotonic (panel D) from constant relative treatment effects. Panels on the right side present the true relations between baseline risk (x-axis) and absolute treatment benefit (y-axis). The 2.5, 25, 50, 75, and 97.5 percentiles of the risk distribution are expressed by the boxplot.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_moderate_base.tiff")))
# knitr::include_graphics(here::here("figures/rmse_base.png"))
```

>>>>

```{r rmsebase, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated from a simulated super-population of size 500,000. The scenario with true constant relative treatment effect (panel A) had a true prediction AUC of 0.75 and sample size of 4250. The RMSE is also presented for strong linear (panel B), strong quadratic (panel C), and non-monotonic (panel D) from constant relative treatment effects. Panels on the right side present the true relations between baseline risk (x-axis) and absolute treatment benefit (y-axis). The 2.5, 25, 50, 75, and 97.5 percentiles of the risk distribution are expressed by the boxplot on the top. The 2.5, 25, 50, 75, and 97.5 percentiles of the true benefit distributions are expressed by the boxplots on the side of the right-handside panel.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_moderate_base.tiff")))
# knitr::include_graphics(here::here("figures/rmse_base.png"))
```

### 2022-02-03 submission/manuscript.tex
new updates to manuscript

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/rmsebase-1} \caption{RMSE of the considered methods across 500 replications calculated from a simulated super-population of size 500,000. The scenario with true constant relative treatment effect (panel A) had a true prediction AUC of 0.75 and sample size of 4250. The RMSE is also presented for strong linear (panel B), strong quadratic (panel C), and non-monotonic (panel D) from constant relative treatment effects. Panels on the right side present the true relations between baseline risk (x-axis) and absolute treatment benefit (y-axis). The 2.5, 25, 50, 75, and 97.5 percentiles of the risk distribution are expressed by the boxplot.}\label{fig:rmsebase}
\end{figure}

>>>>

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/rmsebase-1} \caption{RMSE of the considered methods across 500 replications calculated from a simulated super-population of size 500,000. The scenario with true constant relative treatment effect (panel A) had a true prediction AUC of 0.75 and sample size of 4250. The RMSE is also presented for strong linear (panel B), strong quadratic (panel C), and non-monotonic (panel D) from constant relative treatment effects. Panels on the right side present the true relations between baseline risk (x-axis) and absolute treatment benefit (y-axis). The 2.5, 25, 50, 75, and 97.5 percentiles of the risk distribution are expressed by the boxplot on the top. The 2.5, 25, 50, 75, and 97.5 percentiles of the true benefit distributions are expressed by the boxplots on the side of the right-handside panel.}\label{fig:rmsebase}
\end{figure}

### 2022-02-03 submission/supplement.rmd
new updates to manuscript

\beginsupplement

@@

We set the true risk model coefficients to be $\bm{\beta} = \big(-1.63, 0.26,\dots,0.26\big)^t$ for $AUC=0.65$ and $\bm{\beta} = \big(-2.7, 0.82,\dots,0.82\big)^t$ for $AUC=0.85$. In both cases, $\beta_0$ is selected so that an event rate of $20\%$ is maintained in the control arm.

@@

### Constant treatment effect
Assuming a constant relative treatment effect, fit the adjusted model
in \@ref(eq:risk). Then, an estimate of absolute benefit can be
derived from$$\hat{f}_{\text{benefit}}(lp\given \bm{x},
\hat{\bm{\beta}}) = g(lp) - g(lp+\hat{\gamma}) $$

@@

### Restricted cubic splines
Finally, we drop the linearity assumption and predict absolute benefit using
smoothing with restricted cubic splines with 3, 4, and 5 knots. More
specifically, we fit the model:

$$P(y=1\given lp, t_x) = g\big(\beta_0+\beta_{t_x}t_x+f_{RCS}(lp) + f_{RCS}(lp)\times t_x\big)$$
where
$$f_{RCS}(x)=\alpha_0+\alpha_1h_1(x)+\alpha_2h_2(x)+\dots+\alpha_{k-1}h_{k-1}(x)$$ with
$$h_{j+1}(x)= (x-t_j)^3-(x-t_{k-1})_+\frac{t_k-t_j}{t_k-t_{k-1}}+(x-t_k)^3_+\frac{t_{k-1}-t_j}{t_k-t_{k-   1}}$$
and $t_1,\dots,t_k$ are the selected knots. We predict absolute benefit from $$\hat{f}_{\text{benefit}}(lp \given \bm{x},\hat{\bm{\beta}})=P(y=1\given lp, t_x=0)-P(y=1\given lp,t_x=1)$$.

@@

>>>>

\beginsupplement
\newpage

@@

We set the true risk model coefficients to be
$\bm{\beta} = \big(-1.63,0.26,\dots,0.26\big)^t$ for $AUC=0.65$ and $\bm{\beta} = \big(-2.7,0.82,\dots,0.82\big)^t$ 
for $AUC=0.85$. In both cases, $\beta_0$ is selected so that an event rate of
$20\%$ is maintained in the control arm.

@@

### Constant treatment effect
Assuming a constant relative treatment effect, fit the adjusted model in
\@ref(eq:risk). Then, an estimate of absolute benefit can be derived
from
$$\hat{f}_{\text{benefit}}(lp\given \bm{x}, \hat{\bm{\beta}}) = g(lp) - g(lp+\hat{\gamma}) $$

@@

### Restricted cubic splines
Finally, we drop the linearity assumption and predict absolute benefit using
smoothing with restricted cubic splines with $k=3, 4$ and $5$ knots. More
specifically, we fit the model:

$$P(y=1\given lp, t_x) = g\big(\beta_0+\beta_{t_x}t_x+f_{RCS}(lp) + f_{RCS}(lp)\times t_x\big)$$
where
$$f_{RCS}(x)=\alpha_0+\alpha_1h_1(x)+\alpha_2h_2(x)+\dots+\alpha_{k-1}h_{k-1}(x)$$
with $h_1(x)=x$ and for $j=2,\dots,k-2$
$$h_{j+1}(x)= (x-t_j)^3-(x-t_{k-1})_+^3 \frac{t_k-t_j}{t_k-t_{k-1}}+(x-t_k)^3_+\frac{t_{k-1}-t_j}{t_k-t_{k-1}}$$
where 
$t_1,\dots,t_k$ are the selected knots. We predict absolute benefit from $$\hat{f}_{\text{benefit}}(lp \given \bm{x},\hat{\bm{\beta}})=P(y=1\given lp, t_x=0)-P(y=1\given lp,t_x=1)$$.

@@

\newpage
# Plausible scenario settings

In this section we present specific scenarios from our simulation settings
in which evolution of benefit followed similar patterns to (REF!!). In this case
patients were stratified into risk quarters based on their true baseline
risk. Within each risk quarter we constructed boxplots of true benefit.

```{r, cache=TRUE, echo=FALSE, fig.cap="", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/scenario_397.tiff")))
```

```{r, cache=TRUE, echo=FALSE, fig.cap="", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/scenario_251.tiff")))
```

```{r, cache=TRUE, echo=FALSE, fig.cap="", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/scenario_406.tiff")))
```

```{r, cache=TRUE, echo=FALSE, fig.cap="", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/scenario_324.tiff")))
```

```{r, cache=TRUE, echo=FALSE, fig.cap="", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/scenario_422.tiff")))
```
